---
title: "Forces on a bridge"
author: "Simon Hickinbotham"
date: "29/03/2021"
output:
  html_document: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r obst1, fig.height=4, fig.width=8, message=FALSE, animation.hook="gifski", cache=T, include=FALSE, interval=0.2}
source("maycode.R")

nxl <- c(0,500)
net <- makehexnet(xlim=nxl)
net <- setnodetype(net,50,-5,100,5)
net <- setnodetype(net,410,-5,460,5)
##


inhib=data.frame(x0=190,y0=-10,x1=310,y1=150)

#net <- setnodetype(net,inhib$x0,inhib$y0,inhib$x1,inhib$y1,"X")
net <- setnodetype(net,x0=inhib$x0,y0=inhib$y0,x1=inhib$x1,y1=inhib$y1,"X")

xnet <- net

stim <- data.frame(x=250,y=175)

#plotnet(xnet,stim,inhib,xlim = nxl)

nl <- list()
nl[[1]]<-net

#nl = rungrowth(nl, stim = stim,inhib=inhib, nsteps = 5,xlim=nxl,ylim=c(0,200))

saveRDS(nl,file="ntest1.RDS")

```


```{r gfk, echo=FALSE, message=FALSE, animation.hook="gifski", cache=T, interval=0.2}


source("maycode.R")
net <- xnet
sf <- 100 # scale factor

net$n$x = net$n$x/sf
net$n$y = net$n$y/sf

sinhib <- inhib/sf

sstim <- stim/sf

plotnet(net,sstim,sinhib,xlim = range(net$n$x),ylim=range(net$n$y))

nl <- list()
nl[[1]]<-net

nl = rungrowth(nl, stim = sstim,inhib=sinhib,odist=0.2, nsteps = 25,xlim=range(net$n$x),ylim=range(net$n$y))

source("sodaphysics.R")
subnet <- nl[[length(nl)]]
net<- nl[[length(nl)]]


subnet <- setnodetype(subnet,0.5,-.05,1.1,0.05)
subnet <- setnodetype(subnet,3.9,-0.05,4.6,0.05)
for(ff in 1:1)
plotnet(subnet,sstim,sinhib,xlim = range(net$n$x),ylim=range(net$n$y))


subnet$n = subnet$n[subnet$n$type=="S" | subnet$n$type=="F" ,]
subnet$e = NULL
sei = 1
for(ee in 1:nrow(net$e)){
  if(net$e$from[ee] %in% subnet$n$n & net$e$to[ee] %in% subnet$n$n){
    if(sei == 1)
      subnet$e <- net$e[ee,]
    else
      subnet$e = rbind(subnet$e, net$e[ee,])
    sei = sei + 1
  }
}

for(ff in 1:1)
plotnet(subnet,xlim = range(net$n$x),ylim=range(net$n$y))

gn <- subnet

gn$n$fx <- 0
gn$n$fy <- 0

gn$n$vx <- 0
gn$n$vy <- 0

gn$n$ax <- 0
gn$n$ay <- 0

gn$e$fMag <- 0

#gn$n$y <- gn$n$y + 2
ff <- 0.137
kk <- 6.25
dt <-  0.001666667
ni <- 300

restlength <- VECTOR.len(subnet$n,subnet$e$from[1],subnet$e$to[1])

gnl <- list()

for(tt in 1:50){
  
  
  
  
  #for(dd in 1:300){
  #  gn$n <- clearForce(gn$n)
  #  gn$n <- gForce(gn$n)
  #  gn$n <- fForce(gn$n)#,ff = 0.8)
  #  gn <- kForce(gn)#,kk=20)# Do higher values of kk make the structure 'stiffer'?
  #  gn$n <- integrate(gn$n,dt)
  #}
  gn <- updateModel(gn,ff,kk,dt,ni)
  gnl[[tt]] <- gn

  plotnet(gn,xlim = range(net$n$x),ylim=range(net$n$y))

  nodes <- gn$n
  edges <- gn$e
  for(ee in 1:nrow(gn$e)){
    if(abs(gn$e$fMag[ee])>0.000000000001){
      #message(sprintf("gn$e$fMag[%d] = %0.5f",ee,gn$e$fMag[ee]))
      if(gn$e$fMag[ee]>0){
        n0 <- nodes[nodes$n == edges$from[ee],]
        n1 <- nodes[nodes$n == edges$to[ee],]
        segments(x0=n0$x,x1=n1$x,y0=n0$y,y1=n1$y,lwd = 3,col="red")  
      }else{
        n0 <- nodes[nodes$n == edges$from[ee],]
        n1 <- nodes[nodes$n == edges$to[ee],]
        segments(x0=n0$x,x1=n1$x,y0=n0$y,y1=n1$y,lwd = 3,col="blue")  
      }
    }
      
  }
}
```

ok, Let's see what the forces are doing for the first ten or so frames - are they any use? 


```{r gfkf, echo=FALSE, message=FALSE, animation.hook="gifski", cache=T, interval=0.2}
gn1 <- gnl[[1]]


for(gg in 1:length(gnl)){

  gn <- gnl[[gg]]
  
  #plotnet(gn,xlim = range(net$n$x),ylim=range(net$n$y))
  
  plot(NA,xlim = range(net$n$x),ylim=range(net$n$y))
  
  for(nn in 1:nrow(gn$n)){
    x1 <- gn1$n$x[nn] + (10*gn$n$fx[nn])
    y1 <- gn1$n$y[nn] + (10*gn$n$fy[nn])
    segments(x0=gn1$n$x[nn],y0=gn1$n$y[nn],x1=x1,y1=y1,lwd=4)
    
    x1 <- gn1$n$x[nn] + (10*gn$n$vx[nn])
    y1 <- gn1$n$y[nn] + (10*gn$n$vy[nn])
    segments(x0=gn1$n$x[nn],y0=gn1$n$y[nn],x1=x1,y1=y1,col="red")
    
    x1 <- gn1$n$x[nn] + (10*gn$n$ax[nn])
    y1 <- gn1$n$y[nn] + (10*gn$n$ay[nn])
    segments(x0=gn1$n$x[nn],y0=gn1$n$y[nn],x1=x1,y1=y1,col="green")
    
    
  }
}


```


# Searching over ff and kk

Now we need to find a way of translating these forces into growth responses. A way into this is to look at how the model responds with different values of ff and kk - we want to see if there's a way of having a *fixed* shape, but one which measures *local* forces across the structure via this method. 

To do this, we'll need to put the code above into a function that we can call repeatedly, turning the plot on and off as we go. So let's do that in the R file and call it here: 


```{r ksens, eval=T, include=T}
source("sodaphysics.R")
ff <- 0.137
#kk <- 6.25
dt <-  0.001666667
ni <- 300
for(kk in seq(6,7,1)){
  gn <- subnet
  
  gn$n$fx <- 0
  gn$n$fy <- 0
  gn$n$vx <- 0
  gn$n$vy <- 0
  gn$n$ax <- 0
  gn$n$ay <- 0
  gn$e$fMag <- 0
  
  for(tt in 1:20){
    gn<-updateModel(gn,ff,kk,dt,ni)
  }

  plotnet(gn,xlim = range(net$n$x),ylim=range(net$n$y))
  plotForces(gn)
}



```

```{r}

  plotnet(gn,xlim = range(net$n$x),ylim=range(net$n$y))
plotForces(gn)
saveRDS(gn,file="gndbg.RDS")
```






