---
title: "Experiments in Generative Evolution"
author: "Simon Hickinbotham"
date: "12/01/2021"
output:
  html_document:
    code_folding: hide
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Terminology/strategy

- grow a network of "undifferentiated cells" as "undifferentiated nodes" (u-nodes) connected by "undifferenitated edges" (u-edges) 
- each iteration, check for proximity to a stimulus
- once a stimulus is found, propagate the info back to the foundation
- reinforce edges that make this link
- reposition edges that make this link to support them

## Setting up

Ideally we'd  try to grow from a single foundation, and have a "stimulus point" that triggers conversion from sensing nodes to support nodes. But since determining where to grow - even randomly - is complicated, we'll start with a uniform grid of cells. This is potentially wasteful but it allows us to get to solving the main issues early. 


## Node types

- "U": undifferentiated - all nodes start like this
- "F": foundation - build from here - these are set manually after initialising
- "S": a node near to or linked to a stimulus - which can be a set point or another network
- "X": an undifferentiated node that is located in an inhibition region
- "C": a constructor node - basis of converting to solid structure

## Regular network version

Adding nodes on a grid is pretty straghtforward. Here we'll create a regular square lattice with cross-braces. We commence with the nodes: 
```{r}
source("maycode.R")
```



Now the edges. Let's start by using integers only, cos then we can predict the position and we don't have to minimise any distances to find neighbours. 
Here's a function `addedge` to add an edge to the set:


Now we can generate a regular network within and x,y range - the process will be similar for 3D - we'd just have to add a z dimension. 

```{r}

# constants
iel = 50            # ideal edge length - not used yet
xlim = c(-100,100)  # x range
ylim = c(0,200)     # y range
ostep = 20

net <- makenet(xlim,ylim,ostep)

# Set the foundation nodes: 
#net$n$type[net$n$n>=5 & net$n$n <=7]<-"F"
net <- setnodetype(net,-25,-5,25,5,"F")
 
# Store the network's initial state:
innet <- net

# Here's a stimulus: 
stim <- data.frame(x=-5,y=145)

```


If we eventually want an irregular network then an alternative to the above would be to generate a set of random nodes in the space and then connect the edges via e.g. a Delaunay triangulation. Hard to say what the advantage of this would be right now, but later on it may be a better way to initialise so we can get the system robust to stochastic effects. 


## Growing the network levels


Here's the initial network: a foundation of two connected nodes

Let's have a function to plot the network - we'll use this to keep track of how the network grows later. The bottom of this code block shows the initial network: 

```{r fig.height=6, fig.width=6}
plotnet(net,stim,xlim=xlim,ylim=ylim)
```




```{r include=FALSE}
require(gifski)
```


OK that's a good start - we need to propagate the connectivity between the foundation and the stimulus 

```{r linkit2, animation.hook="gifski", interval = 0.2, fig.width=6, fig.height=6, message=F}

nl <- list()
nl[[1]]<-innet

nl = rungrowth(nl, stim = stim,inhib=NULL, nsteps = 15,xlim=xlim,ylim=ylim)
```

## Other stimulus examples: 

let's do a stimulus in various different positions. Because the lattice is not symmetrical, there'll be some positional effects - again, let's worry about those later when we've got the main network growing. Here's a stimulus to the left of the foundation, which has to run perpendicular to the edge diagonals

```{r linkit3, echo=FALSE, fig.height=6, fig.width=6, message=FALSE, cache=F, animation.hook="gifski", interval=0.2}
stim2 <- stim
stim2$x <- -90

nl <- list()
nl[[1]]<-net

nl = rungrowth(nl, stim = stim2,inhib=NULL, nsteps = 20,xlim=xlim,ylim=ylim)
```

...and here's one to the right, which has to run parallel to the diagonals. You can see there's a different effect on how the network grows. 

```{r linkit, echo=FALSE, fig.height=6, fig.width=6, message=FALSE, cache=F, animation.hook="gifski", interval=0.2}
stim2 <- stim
stim2$x <- 90

nl <- list()
nl[[1]]<-innet

nl = rungrowth(nl, stim = stim2,inhib=NULL, nsteps = 20,xlim=xlim,ylim=ylim)

#net = rungrowth(net = innet, stim = stim2, nsteps = 20, xlim=xlim,ylim=ylim)
```

# Obstacles 

OK, now to start to hit the bridge challenge, we need to specify the void under the bridge. We'll do this via a list of rectangles - this will make it easier to code for in this proof-of-concept stage. I've updated the `plotnet` function to draw obstacles in dark grey behind the network. The next code chunk creates a wider network with a void region, and plots it: 

```{r obst1, cache = F, echo=FALSE, fig.height=4, fig.width=8, message=FALSE, animation.hook="gifski", interval=0.2}
nxl <- c(0,500)
net <- makenet(xlim=nxl)
net <- setnodetype(net,75,-5,125,5)
##


inhib=data.frame(x0=190,y0=-10,x1=310,y1=150)

#net <- setnodetype(net,inhib$x0,inhib$y0,inhib$x1,inhib$y1,"X")
net <- setnodetype(net,x0=inhib$x0,y0=inhib$y0,x1=inhib$x1,y1=inhib$y1,"X")

xnet <- net

stim$x<-410
plotnet(xnet,stim,inhib,xlim = nxl)

nl <- list()
nl[[1]]<-net

nl = rungrowth(nl, stim = stim,inhib=inhib, nsteps = 40,xlim=nxl,ylim=ylim)
```




```{r obst2, cache = F, echo=FALSE, fig.height=5, fig.width=8, message=FALSE, animation.hook="gifski", interval=0.2}

source("maycode.R")
nxl <- c(0,500)
net <- makenet(xlim=nxl)
net <- setnodetype(net,375,-5,425,5)


inhib=data.frame(x0=190,y0=-10,x1=310,y1=150)
net <- setnodetype(net,x0=inhib$x0,y0=inhib$y0,x1=inhib$x1,y1=inhib$y1,"X")

xnet <- net

stim$x<-110
plotnet(xnet,stim,inhib,xlim = nxl)

nl <- list()
nl[[1]]<-net

nl = rungrowth(nl, stim = stim,inhib=inhib, nsteps = 40,xlim=nxl,ylim=ylim)
```







# Multiple networks and stimuli

A key concept is that a node from another network should work exactly like a stimulus. The issue then is how to keep track of stimlevels etc.

Let's refactor everything so we have a bunch of different networks and see how they meet each other around obstacles. Then we can understand what's going on. 


```{r, cache = F, echo=FALSE, fig.height=8, fig.width=12, message=FALSE, animation.hook="gifski", interval=0.2}
source("maycode.R")
nxl <- c(0,500)
net1 <- makenet(xlim=nxl)
net1 <- setnodetype(net1,375,-5,425,5)

inhib=data.frame(x0=190,y0=-10,x1=310,y1=150)
net1 <- setnodetype(net1,x0=inhib$x0,y0=inhib$y0,x1=inhib$x1,y1=inhib$y1,"X")


net2 <- makenet(xlim=nxl)
net2 <- setnodetype(net2,75,-5,125,5)
net2$n$x <- net2$n$x + 10
net2$n$y <- net2$n$y + 10
net2 <- setnodetype(net2,x0=inhib$x0,y0=inhib$y0,x1=inhib$x1,y1=inhib$y1,"X")

netlist<-list()
netlist[[1]]<-net1
netlist[[2]]<-net2


plotnet(net1,NULL,NULL,xlim=nxl)
plotnet(net2,NULL,NULL,xlim=nxl,newplot = F)

netlist = rungrowth(netlist, stim = stim,inhib=inhib, nsteps = 50,xlim=nxl,ylim=ylim)

```

OK, this *kind of* works, but the code is getting unwieldy and there's a lot of redundancy in the way the networks overlap. I can also forsee problems in joining the networks together if we go down this route. 

I think a better approach is to have a *single* network of nodes, but to allow for more than one instance of a node type. So each node has a type (say "F"), and a type index - so we can have more than one Foundation node cluster - in the above image we'd have tidx of 1 for the left side and 2 for the right side. As the levels grow out, we'd keep track of which "F" the level pertains to. What remains to be discovered is how

--------
# How does evolution fit into this

I think we are heading for the situation where each node is governed by a kind of gene-regulatory network, whereby each node updates its state based on information coming from the edges that connect to it and the switching on and off of genes to generate responses. With this in mind, we should consider changing the processing of the data to a node-by-node basis. So let's do the following:

- move to a network built from equilateral triangles
- accommodate multiple foundations and stimuli in a single network
- emphasise processing on a per-node basis

On this basis, I think it's time to start a new workbook and implement these ideas. 

--------
# Obsolete code

Here are some functions I tried but that didn't go anywhere (yet)

#### Growing the network from scratch

I did a little work on actually growing the network, rather than establishing a grid that covers the whole space. 

Here's the initial network: a foundation of two connected nodes

```{r}
# constants
iel = 50

nodes <- data.frame(n = c(0,1),x = c(-25,25),y=c(0,0),level=c(0,0),type=c("F","F"),e1=c(F,F),e2=c(F,F),e3=c(T,F),e4=c(F,F),e5=c(F,F),e6=c(F,T))
edges <- data.frame(from = 0, to = 1, type = "F")
net <- list()
net$n <- nodes
net$e <- edges
```

#### Edge numbering

If we need it...

```{r echo=FALSE, fig.height=4, fig.width=4}
plot(NA,xlim=c(-30,30),ylim=c(-30,30),asp=T,xlab="x",ylab="y")

# Do the edges first
segments(x0=0,y0=0,x1=  0,y1= 25)
segments(x0=0,y0=0,x1= 25,y1= 25)
segments(x0=0,y0=0,x1= 25,y1=  0)
segments(x0=0,y0=0,x1=  0,y1=-25)
segments(x0=0,y0=0,x1=-25,y1=-25)
segments(x0=0,y0=0,x1=-25,y1=  0)

# Now the nodes
points(x=0,y=0,pch=21,col="black",bg="black",cex=2)

points(x=  0,y= 25,pch=21,col="black",bg="white",cex=2)
points(x= 25,y= 25,pch=21,col="black",bg="white",cex=2)
points(x= 25,y=  0,pch=21,col="black",bg="white",cex=2)
points(x=  0,y=-25,pch=21,col="black",bg="white",cex=2)
points(x=-25,y=-25,pch=21,col="black",bg="white",cex=2)
points(x=-25,y=  0,pch=21,col="black",bg="white",cex=2)

# index
text("1",x=  2,y= 20)
text("2",x= 20,y= 17)
text("3",x= 20,y=-2)
text("4",x=-2 ,y=-20)
text("5",x=-20,y=-17)
text("6",x=-20,y= 2)
```