---
title: "Experiments in Generative Evolution"
author: "Simon Hickinbotham"
date: "12/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Terminology/strategy

- grow a network of "undifferentiated cells" as "undifferentiated nodes" (u-nodes) connected by "undifferenitated edges" (u-edges) 


# Setting up

Ideally we'd  try to grow from a single foundation, and have a "stimulus point" that triggers conversion from sensing nodes to support nodes. But since determining where to grow - even randomly - is complicated, we'll start with a uniform grid of cells. This is potentially wasteful but it allows us to get to solving the main issues early. 

To check how the logic might work in a real space, let's set up some assumptions. We'll store the functions/classes in the following code block as we develop until we've got that far. 

## Edge numbering

```{r echo=FALSE, fig.height=4, fig.width=4}
plot(NA,xlim=c(-30,30),ylim=c(-30,30),asp=T,xlab="x",ylab="y")

# Do the edges first
segments(x0=0,y0=0,x1=  0,y1= 25)
segments(x0=0,y0=0,x1= 25,y1= 25)
segments(x0=0,y0=0,x1= 25,y1=  0)
segments(x0=0,y0=0,x1=  0,y1=-25)
segments(x0=0,y0=0,x1=-25,y1=-25)
segments(x0=0,y0=0,x1=-25,y1=  0)

# Now the nodes
points(x=0,y=0,pch=21,col="black",bg="black",cex=2)

points(x=  0,y= 25,pch=21,col="black",bg="white",cex=2)
points(x= 25,y= 25,pch=21,col="black",bg="white",cex=2)
points(x= 25,y=  0,pch=21,col="black",bg="white",cex=2)
points(x=  0,y=-25,pch=21,col="black",bg="white",cex=2)
points(x=-25,y=-25,pch=21,col="black",bg="white",cex=2)
points(x=-25,y=  0,pch=21,col="black",bg="white",cex=2)

# index
text("1",x=  2,y= 20)
text("2",x= 20,y= 17)
text("3",x= 20,y=-2)
text("4",x=-2 ,y=-20)
text("5",x=-20,y=-17)
text("6",x=-20,y= 2)
```



# Growing the network

Here's the initial network: a foundation of two connected nodes

```{r}
# constants
iel = 50

nodes <- data.frame(n = c(0,1),x = c(-25,25),y=c(0,0),level=c(0,0),type=c("F","F"),e1=c(F,F),e2=c(F,F),e3=c(T,F),e4=c(F,F),e5=c(F,F),e6=c(F,T))
edges <- data.frame(from = 0, to = 1, type = "F")
net <- list()
net$n <- nodes
net$e <- edges
```

Let's have a function to plot the network - we'll animate this later: 

```{r}
plotnet <- function(net){
  nodes <- net$n
  edges <- net$e
  plot(NA,xlim=c(-100,100),ylim=c(0,500),asp=T)
  
  #Draw the edges
  for(ee in 1:nrow(edges)){
    n0 <- nodes[nodes$n == edges$from[ee],]
    n1 <- nodes[nodes$n == edges$to[ee],]
    segments(x0=n0$x,x1=n1$x,y0=n0$y,y1=n1$y,lwd = 2,col="black")
  }
  
  
  #Draw the nodes
  points(x=nodes$x,y=nodes$y,pch=20,cex=2)
}

plotnet(net)
```

Now we need a function to propagate the network


```{r}
verbose = T
#propagate <- function(net,verbose=T){
  #randomly pick a node
 node <- net$n[sample(nrow(net$n),1),]
  if(verbose)message(sprintf("Chose node %d", node$n))
  
  #randomly pick a neighbour
  edges <- net$e[net$e$from == node$n | net$e$to == node$n,]
  if(verbose)message(sprintf("Edges found =  %d", nrow(edges)))
  
  #add a node in new space - let's be naive and put it between one of the edges
  edge <- edges[sample(nrow(edges),1),]
  
  #now get the partner node
  if(edge$from == node$n){
    n2 <- net$n[net$n$n == edge$to,]
  }else{
    n2 <-  net$n[net$n$n == edge$from,]
  }
  
  newx <- (node$x + n2$x)/2
  newy <- (node$y + n2$y)/2
  
  #add edges
  
#}

propagate(net)
```



# Tests


```{r}

```



