---
title: "Version2"
author: "Simon Hickinbotham"
date: "16/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

We plan to:

- move to a network built from equilateral triangles
- accommodate multiple foundations and stimuli in a single network
- emphasise processing on a per-node basis

On this basis, I think it's time to start a new workbook and implement these ideas. 

# Setting up the network

## Node types

- "U": undifferentiated - all nodes start like this
- "F": foundation - build from here - these are set manually after initialising
- "S": a node near to or linked to a stimulus - which can be a set point or another network
- "X": an undifferentiated node that is located in an inhibition region
- "C": a constructor node - basis of converting to solid structure

## Regular hexagonal network version

Adding nodes on a grid is pretty straghtforward. Here we'll create a regular square lattice with cross-braces. We commence with the nodes: 
```{r}
source("maycode.R")
```


Now we can generate a regular network within and x,y range - the process will be similar for 3D - we'd just have to add a z dimension. 

```{r fig.height=6, fig.width=6, cache=TRUE}
source("maycode.R")

# constants
iel = 50            # ideal edge length - not used yet
xlim = c(-100,100)  # x range
ylim = c(0,200)     # y range
ostep = 20

net <- makehexnet(xlim,ylim,ostep)

# Set the foundation nodes: 
#net$n$type[net$n$n>=5 & net$n$n <=7]<-"F"
net <- setnodetype(net,-25,-5,25,5,"F")
 
# Store the network's initial state:
innet <- net

# Here's a stimulus: 
stim <- data.frame(x=-5,y=145)


plotnet(net,stim)
```



```{r include=FALSE}
require(gifski)
```


OK that's a good start - we need to propagate the connectivity between the foundation and the stimulus 

```{r linkit2, animation.hook="gifski", interval = 0.2, fig.width=6, fig.height=6, message=F, cache=T}

nl <- list()
nl[[1]]<-innet

nl = rungrowth(nl, stim = stim,inhib=NULL, nsteps = 15,xlim=xlim,ylim=ylim)

```



```{r linkit3, echo=FALSE, fig.height=6, fig.width=6, message=FALSE, cache=T, animation.hook="gifski", interval=0.2}
stim2 <- stim
stim2$x <- -90

nl <- list()
nl[[1]]<-net

nl = rungrowth(nl, stim = stim2,inhib=NULL, nsteps = 20,xlim=xlim,ylim=ylim)

#net = rungrowth(net = innet, stim = stim2, nsteps = 20, xlim=xlim,ylim=ylim)
```
```{r linkit, echo=FALSE, fig.height=6, fig.width=6, message=FALSE, cache=F, animation.hook="gifski", interval=0.2}
stim2 <- stim
stim2$x <- 90

nl <- list()
nl[[1]]<-innet

nl = rungrowth(nl, stim = stim2,inhib=NULL, nsteps = 20,xlim=xlim,ylim=ylim)
```

# Obstacles 

OK, now to start to hit the bridge challenge, we need to specify the void under the bridge. We'll do this via a list of rectangles - this will make it easier to code for in this proof-of-concept stage. I've updated the `plotnet` function to draw obstacles in dark grey behind the network. The next code chunk creates a wider network with a void region, and plots it: 

```{r obst1, cache = T, echo=FALSE, fig.height=4, fig.width=8, message=FALSE, animation.hook="gifski", interval=0.2}
nxl <- c(0,500)
net <- makehexnet(xlim=nxl)
net <- setnodetype(net,75,-5,125,5)
##


inhib=data.frame(x0=190,y0=-10,x1=310,y1=150)

#net <- setnodetype(net,inhib$x0,inhib$y0,inhib$x1,inhib$y1,"X")
net <- setnodetype(net,x0=inhib$x0,y0=inhib$y0,x1=inhib$x1,y1=inhib$y1,"X")

xnet <- net

stim$x<-410
plotnet(xnet,stim,inhib,xlim = nxl)

nl <- list()
nl[[1]]<-net

nl = rungrowth(nl, stim = stim,inhib=inhib, nsteps = 40,xlim=nxl,ylim=ylim)

saveRDS(nl,file="ntest1.RDS")

```


# Calculating forces - sodaracer version

Let's look at how sodaracer-type physics engines sorts out forces in their meshes - perhaps not an obvious choice but may provide a way 'in' to the system. 

A 'model' in soda racer consists of a bunch of masses (nodes), connected by springs (edges) - very similar to the model we are using above. 

Each mass has parameters {s,v(velocity),a(acceleration),f (force)}, each of which has an xy coordinate. 

Each spring has 2 node names, restlength, amplitude, phase

Each iteration updates the masses. 6 steps: clear; gForce(gravity); fForce; kForce; integrate; collide. 

- clearForces sets the f vector to zero - so the starting f values in the model are immediately deleted
- gForce simply adds a fixed unit to the y-dim of 'f'
- fForce - calculates friction (opposes velocity) - a vector multiplication of the f and v vectors
- kForce - apply springwise force to both masses in the spring
- integrate. F = ma, so a = F/m. 
- collide checks if a mass has hit a wall - we don't need this!


It looks like this approach uses Euler's method to update the forces over small delta-time. 



