---
title: 'Evolvable Shape Seeds: Progress at York'
author: "Simon Hickinbotham"
date: "4 December 2019"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



## Summary

- Algorithm for 'growing' a set of seed genes
- Strengths and weaknesses
- Comparision with L-Systems and CPPNs
- Novelty search

## Target Shape


```{r}
summary(cars)
```

## Algorithm pseudocode

<small>
```{r, echo=TRUE, eval = FALSE}
genetostlfile4 <- function(fn="shape.stl",genes,pos=spos(0,0,0),runlim=1000){

  while (TRUE)
    posnext <- NULL
    for(pp in 1:nrow(pos)){
      
      shape <- getshape(pos[pp,],genes)
      domval <- domInZone(pos[pp,],genes)
      if(domval){
        
        dirs <- genes[genes$dom == domval & dirs$valtyp!=0,] # extract domlevel
        for(dd in 1:nrow(dirs)){
          direction <- get_direction(dirs)
          posnext <- addposition(dirs[dd],shape,direction,posnext)
          drawCStoSTL(pos[pp,],shape,direction)
    
    if(is.null(posnext) | iterations++ > runlim)      
      break
    
    pos <- posnext
```
</small>


## Seed genes



novelty search as a means of measuring the "expressive power" of an encoding



## Strengths & Weaknesses


- Very flexible *and readable*
- Compatible with engineering platforms

- Arbitrary start/stop points
- Not particularly compact
- X,Y,Z zones are only loosely joined

