---
title: "Novelty Search"
author: "Simon Hickinbotham"
date: "29/11/2019"

output: 
  html_document: 
    number_sections: yes
  pdf_document: default
---

# Shape Seeds

## Table shape seed

The seed shape is the "table" described in the file `091111Progress.Rmd`. Here we've just tidied up some of the values and names. The code at the end of this chunk generates the shape from the geneset and gets a list of points from the phenotype:

```{r echo=TRUE, message=FALSE}

require(stringr)
source("../shapevol1/R/sgene.R")
source("../shapevol1/R/genetostl.R")
source("../shapevol1/R/genetostl4.R")

gene01 <- sgene("Cross section","Square",status = T,start=-40,stop=40,dom=1)
gene02 <- sgene("Length",       5       ,status = T,start=-40,stop=40,dom=1)
gene03 <- sgene("Diameter",     1       ,status = T,start=-40,stop=40,dom=0)
group1 <- rbind(gene01,gene02,gene03)

gene04 <- sgene("X_1X",          0,status = T,start=0,stop=1,dom=50)
gene05 <- sgene("Y_1Y",          0,status = T,start=0,stop=1,dom=50)
gene06 <- sgene("Z_1Z",          1,status = T,start=0,stop=5,dom=50)
group2 <- rbind(gene04,gene05,gene06)

#Table top:
gene07 <- sgene("T1_2X",          1,status = T,start=0,stop=5,dom=48)
gene08 <- sgene("T2_2Y",          1,status = T,start=0,stop=10,dom=48)
gene09 <- sgene("T1_2Z",          0,status = T,start=0,stop=15,dom=48)

#y axis leg and end strut 
gene10 <- sgene("T3_2X",          1,status = T,start=0,stop=1,dom=47)
gene11 <- sgene("T3_2Y",          0,status = T,start=15,stop=16,dom=47)
gene12 <- sgene("T3_2Z",          -1,status = T,start=6,stop=10,dom=47)

#x axis leg and end strut
gene13 <- sgene("T4_2X",          0,status = T,start=10,stop=11,dom=46)
gene14 <- sgene("T4_2Y",          1,status = T,start=0,stop=4,dom=46)
gene15 <- sgene("T4_2Z",          -1,status = T,start=6,stop=10,dom=46)

# Table edge along far y axis
gene16 <- sgene("T5_2X",          0,status = T,start=10,stop=11,dom=45)
gene17 <- sgene("T5_2Y",          1,status = T,start=4,stop=14,dom=45)
gene18 <- sgene("T5_2Z",          0,status = T,start=6,stop=10,dom=45)

# Table edge along far x axis
gene19 <- sgene("T6_2X",          1,status = T,start=5,stop=9,dom=44)
gene20 <- sgene("T6_2Y",          0,status = T,start=11,stop=15,dom=44)
gene21 <- sgene("T6_2Z",          0,status = T,start=6,stop=10,dom=44)

# Lower portions of the remaining three legs
gene22 <- sgene("L3_2X",          0,status = T,start= 0,stop= 100,dom=43)
gene23 <- sgene("L3_2Y",          0,status = T,start=0,stop=15,dom=43)
gene24 <- sgene("L3_2Z",          -1,status = T,start= 1,stop=12,dom=43)

group3 <- rbind(gene07,gene08,gene09
                ,gene10,gene11,gene12
                ,gene13,gene14,gene15
                ,gene16,gene17,gene18
                ,gene19,gene20,gene21
                ,gene22,gene23,gene24
                )

geneset <- rbind(group1,group2,group3)

pointset <- genetostlfile4   ("~/Desktop/stl/seedX4g3_00_00_00.stl",geneset,pos=spos(0, 0,0),runlim=40)
```

That makes the table alright. It's nice to generate an STL file with all the steps in for visualisation - let's do that now:


```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE, paged.print=FALSE}
  
fn <- "~/git/cec2020/seedsteps.stl"
sink(file = fn)
cat("solid Exported from Blender-2.80 (sub 75)\n")

xpos <- 200
xs <- -25
ypos <- -50
ys <- 3
geneset <- rbind(group1,group2)
pointset <- genetostlfile4   ("~/Desktop/stl/seedX4g3_00_00_00.stl",geneset,
                              pos=spos(0, 0,0),runlim=40,offset=c(xpos,ypos,0),solo = F)
xpos <- xpos + xs
ypos <- ypos + ys
geneset <- rbind(geneset,gene07,gene08,gene09)
pointset <- genetostlfile4   ("~/Desktop/stl/seedX4g3_00_00_00.stl",geneset,
                              pos=spos(0, 0,0),runlim=40,offset=c(xpos,ypos,0),solo = F)
xpos <- xpos + xs
ypos <- ypos + ys
geneset <- rbind(geneset,gene10,gene11,gene12)
pointset <- genetostlfile4   ("~/Desktop/stl/seedX4g3_00_00_00.stl",geneset,
                              pos=spos(0, 0,0),runlim=40,offset=c(xpos,ypos,0),solo = F)
xpos <- xpos + xs
ypos <- ypos + ys
geneset <- rbind(geneset,gene13,gene14,gene15)
pointset <- genetostlfile4   ("~/Desktop/stl/seedX4g3_00_00_00.stl",geneset,
                              pos=spos(0, 0,0),runlim=40,offset=c(xpos,ypos,0),solo = F)
xpos <- xpos + xs
ypos <- ypos + ys
geneset <- rbind(geneset,gene16,gene17,gene18)
pointset <- genetostlfile4   ("~/Desktop/stl/seedX4g3_00_00_00.stl",geneset,
                              pos=spos(0, 0,0),runlim=40,offset=c(xpos,ypos,0),solo = F)
xpos <- xpos + xs
ypos <- ypos + ys
geneset <- rbind(geneset,gene19,gene20,gene21)
pointset <- genetostlfile4   ("~/Desktop/stl/seedX4g3_00_00_00.stl",geneset,
                              pos=spos(0, 0,0),runlim=40,offset=c(xpos,ypos,0),solo = F)
xpos <- xpos + xs
ypos <- ypos + ys
geneset <- rbind(geneset,gene22,gene23,gene24)
pointset <- genetostlfile4   ("~/Desktop/stl/seedX4g3_00_00_00.stl",geneset,
                              pos=spos(0, 0,0),runlim=40,offset=c(xpos,ypos,0),solo = F)

cat("solid Exported from Blender-2.80 (sub 75)\n")
sink()


```



## Shape 2 - a 'manual' mutation of Shape1

This shape is interesting because it raised the need to include *all* the 'already seen' positions in the 'addpos' functions. This is becuase each 'nextpos' structure didn't contain duplicates, but over several iterations, positions were revisited many times - indicating a 'loop' of positions that can only be detected if we use the 'seenpos' structure to record every position that has already been processed. If we don't do this, we end up hitting the iteration limit, and the full shape isn't produced. 



```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE, paged.print=FALSE}

gene01 <- sgene("Cross section","Square",status = T,start=-40,stop=40,dom=1)
gene02 <- sgene("Length",       5       ,status = T,start=-40,stop=40,dom=1)
gene03 <- sgene("Diameter",     0.5       ,status = T,start=-40,stop=40,dom=0)
group1 <- rbind(gene01,gene02,gene03)

gene04 <- sgene("X_1X",          0,status = T,start=0,stop=1,dom=50)
gene05 <- sgene("Y_1Y",          0,status = T,start=0,stop=1,dom=50)
gene06 <- sgene("Z_1Z",          1,status = T,start=0,stop=5,dom=50)
group2 <- rbind(gene04,gene05,gene06)

#Table top:
gene16 <- sgene("T1_2X",          1,status = T,start=0,stop=5,dom=48)
gene17 <- sgene("T2_2Y",          1,status = T,start=0,stop=10,dom=48)
gene18 <- sgene("T1_2Z",          0,status = T,start=0,stop=15,dom=48)

#y axis leg and end strut 
gene19 <- sgene("T3_2X",          1,status = T,start=0,stop=1,dom=47)
gene20 <- sgene("T3_2Y",          0,status = T,start=15,stop=16,dom=47)
gene21 <- sgene("T3_2Z",          -1,status = T,start=6,stop=10,dom=47)

#x axis
gene22 <- sgene("T4_2X",          1,status = T,start=10,stop=11,dom=46)
gene23 <- sgene("T4_2Y",          1,status = T,start=0,stop=4,dom=46)
gene24 <- sgene("T4_2Z",          -1,status = T,start=6,stop=10,dom=46)


gene25 <- sgene("T5_2X",          0,status = T,start=0,stop=20,dom=45)
gene26 <- sgene("T5_2Y",          1,status = T,start=-4,stop=20,dom=45)
gene27 <- sgene("T5_2Z",          1,status = T,start=-6,stop=20,dom=45)

gene28 <- sgene("T6_2X",          1,status = T,start=5,stop=9,dom=44)
gene29 <- sgene("T6_2Y",          0,status = T,start=11,stop=15,dom=44)
gene30 <- sgene("T6_2Z",          0,status = T,start=6,stop=10,dom=44)

gene31 <- sgene("L3_2X",          0,status = T,start= 0,stop= 100,dom=43)
gene32 <- sgene("L3_2Y",          0,status = T,start=0,stop=15,dom=43)
gene33 <- sgene("L3_2Z",          -1,status = T,start= 1,stop=12,dom=43)

gene34 <- sgene("L4_2X",          1,status = T,start= 10,stop= 11,dom=51)
gene35 <- sgene("L4_2Y",          0,status = T,start=15,stop=16,dom=51)
gene36 <- sgene("L4_2Z",          -1,status = T,start= 1,stop=10,dom=51)

#Legs:
group3 <- rbind(gene16,gene17,gene18
                ,gene19,gene20,gene21
                ,gene22,gene23,gene24
                ,gene25,gene26,gene27
                ,gene28,gene29,gene30
                ,gene31,gene32,gene33
                ,gene34,gene35,gene36
                )


genesetm <- rbind(group1,group2,group3)

require(stringr)
source("../shapevol1/R/sgene.R")
source("../shapevol1/R/genetostl.R")
source("../shapevol1/R/genetostl4.R")
pointsetm <- genetostlfile4   ("~/Desktop/stl/seedX4g3_00_00_00_mutant1.stl",genesetm,pos=spos(0, 0,0),offset = c(10,10,10),runlim=40)
```


# Building the Novelty search

OK - so for novelty search, we need to do two things:

1. Get some novelty metrics from the pointset
2. Get a distance measure from the novelty metrics
2. Mutate the genes. 

We can then begin to measure the novelty as the genes mutate. Let's start by getting some simple stats out of the positions for the novelty metrics:

- X,Y,Z range
- X,Y,Z mean
- Num iterations
- Num positions

We have a function called `getNSvals` that does this for a pointset:

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE, paged.print=FALSE}
source("../shapevol1/R/nsvals.R")
nspos <- getNSvals(pointset)
```

## Mutation function

OK, let's write some functions to help us generate a bunch of mutants from the original table:

*TODO: these should be in the R package*

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE, paged.print=FALSE}

getrowno <- function(gno,xyzno){
  rno <- (3*gno)+xyzno
  return (rno)
}
source("../shapevol1/R/domut.R")

mutate.gene <- function(geneset,verbose=F){

  #select a genegroup for direction:
  gno <- sample(((nrow(geneset)/3)-1), 1)
  xyzno <- sample(3,1)-1
  
  #pick a mutant type
  mtype <- sample(4,1)
  
  mdir <- sample(2,1)
  
  geneset <- domut(gno,xyzno,mtype,mdir,geneset,verbose)
  
  return (geneset)

}


NSdist <- function(v1,v2){
  dist <- 0
  for(vv in 1:ncol(v1)){
    dist <- dist + ((v1[1,vv]-v2[1,vv])*(v1[1,vv]-v2[1,vv]))
  }
  
  dist <- sqrt(dist)
  return(dist)
  
}
```


A crude test of the above functions yields: 


```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE, paged.print=FALSE}
gpop <- list()
ppop <- list()
npop <- list()
dpop <- list()

geneset.points <- genetostlfile4   (sprintf("~/Desktop/stl/pop00.stl"),geneset,pos=spos(0, 0,0),runlim=100)
geneset.ns <- getNSvals(geneset.points)
geneset.dist <- 0

for(gg in 1:25){
  gpop[[gg]] <- mutate.gene(geneset)
  ppop[[gg]] <- genetostlfile4   (sprintf("~/Desktop/stl/pop%02d.stl",gg),gpop[[gg]],pos=spos(0, 0,0),runlim=100)
  npop[[gg]] <- getNSvals(ppop[[gg]])
  dpop[[gg]] <- NSdist(npop[[1]],npop[[gg]])
}
```

## Testing individual mutations

OK, this is now giving us an initial population to work with, so we can look around, but we have a bug to deal with - let's brute force the mutations to find out where it happens, and make sure we can load *all* the mutants into one file:

*(NB! This has strange effects when knitting to a document due to the `sink` function... to get results you have to use CTRL + ENTER on the code block..)*



```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE, paged.print=FALSE}
glim <- 7
xlim <- 3
mlim <- 4

source("../shapevol1/R/genetostl.R")
source("../shapevol1/R/genetostl4.R")

fn <- "mutatedShapes1.stl"  
sink(file = fn)
cat("solid Exported from Blender-2.80 (sub 75)\n")

for(gg in 1:glim){
  for(xx in 1:xlim){
    message(sprintf("rowno = %d gg = %d, xx = %d",getrowno(gg,xx),gg,xx))
    for(mm in 1:mlim){
      g2 <- domut(gg,xx,mm,sample(2,1),geneset,verbose=F)
      pv <- genetostlfile4(fn, g2,offset=c((gg-(glim/2))*30,(xx-(xlim/2))*30,(mm-(mlim/2))*30),verbose = F, runlim = 50,solo=F)
    }
  }
}

cat("solid Exported from Blender-2.80 (sub 75)\n")
sink()
```

OK, that now works nicely so we can put it together into an algorithm. First though, it might be a good idea to see how far 100 random mutations might take us - an alternative view of the 'mutation space'. 



```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE, paged.print=FALSE}
glim <- 7
xlim <- 3
mlim <- 4

source("../shapevol1/R/genetostl.R")
source("../shapevol1/R/genetostl4.R")

fn <- "shapes100mutns.stl"  
sink(file = fn)
cat("solid Exported from Blender-2.80 (sub 75)\n")

#for(gg in 1:glim){
#  for(xx in 1:xlim){
#    message(sprintf("rowno = %d gg = %d, xx = %d",getrowno(gg,xx),gg,xx))
#    for(mm in 1:mlim){
  
pv<-list()    

xpos <- -250
ypos <- -250
for(nn in 1:100){
  message(sprintf("Mutating sample %d 100 times",nn))
  g2 <- geneset
  for(mm in 1:100){
    g2 <- domut(sample(glim,1),sample(xlim,1),sample(mlim,1),sample(2,1),g2,verbose=F)
  }
  pv[[nn]] <- genetostlfile4(fn, g2,offset=c(xpos,ypos,0),verbose = F, runlim = 100,solo=F)
  xpos<-xpos + 50
  if(xpos > 200){
    xpos <- -250
    ypos <- ypos + 50 
  }
}

cat("solid Exported from Blender-2.80 (sub 75)\n")
sink()
```


OK Now we want to get the Novelty Search vectors for each of these shapes: 


```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE, paged.print=FALSE}
nsv <- data.frame(xmean=double(),ymean=double(),zmean=double(),xrange=integer(),yrange=integer(),zrange=integer(),nits=integer(),npos=integer())
for(nn in 1:100)
  nsv[nn,]<-getNSvals(pv[[nn]])

boxplot(nsv$xmean,nsv$ymean,nsv$zmean,nsv$xrange,nsv$yrange,nsv$zrange,nsv$nits,nsv$npos,ylim=c(0,50))
```

# Population to stl file 



```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE, paged.print=FALSE}
pop.to.stl <- function(fn = "population.stl",  glist){

  sink(file = fn)
  cat("solid Exported from Blender-2.80 (sub 75)\n")
  xpos <- -225
  ypos <- -225
  for(pp in 1:length(glist)){
    ppop <- genetostlfile4   (sprintf("tmp.stl"),glist[[pp]],pos=spos(0, 0,0),offset=c(xpos,ypos,0),runlim=100,solo=F)
    xpos<-xpos + 50
    if(xpos == 275){
      xpos <- -225
      ypos <- ypos + 50 
    }
  }
  cat("solid Exported from Blender-2.80 (sub 75)\n")
  sink()
}
```


# Algorithm 1

*simple attempt to get some evolution going and explore the issues*

- Create empty list of 'seen' states
- Add 'table' to this list
- Generate population by mutating the table
- while not finished:
  - use metrics on the population to rank distance from 'seen' states
  - replace bottom $m$ shapes in the ranking with mutants of the top $m$ shapes
  - if distance $d>D$, the minimum distance for a new state to be seen, add that to the 'seen' states



```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE, paged.print=FALSE}

#reload all the functions!
source("../shapevol1/R/domut.R")


#set the random number seed for repeatable results
set.seed(436)

# Create an empty list of seen states
nseen <- 0
seen <- list()
seenNS <- list()
popsize <- 25

gpop <- list()
ppop <- list()
npop <- list()
dpop <- list()

# Add 'table' to this list
seen[[1]] <- geneset
seenNS[[1]] <- getNSvals(genetostlfile4("geneset.stl",geneset))
  
nseen <- nseen + 1
    # message(sprintf("length of seenNS is %d",length(seenNS)))
    # for(sl in 1:length(seenNS)){
    #   dist <-  NSdist(seenNS[[sl]],npop[[ mut ]])
    #   dpop[[  mut  ]] <- min(dpop[[mut]], dist)
    #   message(sprintf("dpop[[mut]] = %f, dist = %f",dpop[[mut]],dist))
    # }



fn <- "gen1.stl"  
sink(file = fn)
cat("solid Exported from Blender-2.80 (sub 75)\n")
xpos <- -100
ypos <- -100
for(pp in 1:popsize){
  gpop[[pp]] <- mutate.gene(geneset)
  ppop[[pp]] <- genetostlfile4   (sprintf("tmp.stl"),gpop[[pp]],pos=spos(0, 0,0),offset=c(xpos,ypos,0),runlim=100,solo=F)
  npop[[pp]] <- getNSvals(ppop[[pp]])
  dpop[[pp]] <- NSdist(npop[[1]],npop[[pp]])
  xpos<-xpos + 50
  if(xpos == 150){
    xpos <- -100
    ypos <- ypos + 50 
  }
}
cat("solid Exported from Blender-2.80 (sub 75)\n")
sink()

nmuts <- 5
distthreshold <- 10
ngen <- 100

for(iterations in 1:ngen){
  message(sprintf("\n\nITERATION %d",iterations))
  
  #TODO: we should only do this if new seen states have been added 
  #update *all* the scores
  for(gg in 1:length(gpop)){
    dpop[[ gg ]] <- NSdist(seenNS[[1]],npop[[ gg ]])
    if(length(seenNS)>1){
      for(ss in 2:length(seenNS)){
        dist <-  NSdist(seenNS[[ss]],npop[[ gg ]])
        dpop[[  gg  ]] <- min(dpop[[gg]], dist)
        
      }
    }
  }
  

  #Generate the mutants and update the scores: 
  myorder <- order(as.data.frame(dpop))
  for(mm in 1:nmuts){
    mut <- myorder[mm]
    msrc <-  myorder[mm+(popsize-nmuts)] 
    message(sprintf("Mutating gene %d with dpop score of %f into gene %d with dpop score of %f",
                    msrc,dpop[[msrc]],mut,dpop[[mut]]))
    gpop[[  mut  ]]   <- mutate.gene(gpop[[ msrc ]])
    ppop[[  mut  ]] <- genetostlfile4   (sprintf("~/Desktop/stl/gen%dpop%02d.stl",iterations,mut),
                                         gpop[[  mut  ]],pos=spos(0, 0,0),runlim=100)
    npop[[  mut  ]] <- getNSvals(ppop[[  mut  ]])
    
    # Get the lowest distance score for this mutant: 
    dpop[[ mut ]] <-  NSdist(seenNS[[1]],npop[[ mut ]])
    # message(sprintf("length of seenNS is %d",length(seenNS)))
    # for(sl in 1:length(seenNS)){
    #   dist <-  NSdist(seenNS[[sl]],npop[[ mut ]])
    #   dpop[[  mut  ]] <- min(dpop[[mut]], dist)
    #   message(sprintf("dpop[[mut]] = %f, dist = %f",dpop[[mut]],dist))
    # }
    
    message(sprintf("dpop[[%d]] is now %f",mut,dpop[[mut]]))
    
  }
  
  #Add high-scoring genesets to the list of seen shapes 
  seenlen <- length(seen)
  #for(dd in 1:popsize){
  
  
  # #Only add one geneset per generation, and only if it's above D
  # # reason: if we added more then one, we'd have to recalculate *all* distances for each new entry in the seen list...
  # myorder <- order(as.data.frame(dpop))
  # maxd <- myorder[length(dpop)]
  # if(dpop[[ maxd ]] > distthreshold){
  #   seenlen <- seenlen + 1
  #   seen[[seenlen]] <- gpop[[maxd]]
  #   seenNS[[seenlen]] <- getNSvals(genetostlfile4("geneset.stl",seen[[seenlen]]))
  #   message(sprintf("Added gene %d from pop with dist of %f, seenlen now %d",dd,dpop[[maxd]],seenlen))
  # }
  
  pop.to.stl(sprintf("seen%02d.stl",iterations),seen)
  pop.to.stl(sprintf("gpop%02d.stl",iterations),gpop)
}


```


We need a function to view the outputs:


```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE, paged.print=FALSE}

pop.to.stl("gpop.stl",gpop)
pop.to.stl("seen.stl",seen)

```



## Evaluation

OK, so the problem with the output of this as it stans is that 




Ways we might vary this:

- Use a 'target' shape as the distance measure - to see if we can mutate between shapes - is the grammar flexible enough?
  
  















