---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---


## Test the start block

```{r echo=TRUE, message=FALSE}

require(stringr)
source("../shapevol1/R/sgene.R")
source("../shapevol1/R/drawCStoSTL.R")
source("../shapevol1/R/genetostl4.R")

gene01 <- sgene("Cross section","Square",status = T,start=-40,stop=40,dom=1)
gene02 <- sgene("Length",       5       ,status = T,start=-40,stop=40,dom=1)
gene03 <- sgene("Diameter",     1       ,status = T,start=-40,stop=40,dom=1)
group1 <- rbind(gene01,gene02,gene03)


pointset <- genetostlfile4   ("~/Desktop/stl/teststart.stl",group1,pos=spos(0, 0,0),runlim=40,verbose = T)
```

## Test shape

The test shape is pretty simple: start at position 0,0,0 and branch out in all six possible directions. Then see which slicer has problems with the result. 

```{r echo=TRUE, message=FALSE}

require(stringr)
source("../shapevol1/R/sgene.R")
source("../shapevol1/R/drawCStoSTL.R")
source("../shapevol1/R/genetostl4.R")

gene01 <- sgene("Cross section","Square",status = T,start=-40,stop=40,dom=1)
gene02 <- sgene("Length",       5       ,status = T,start=-40,stop=40,dom=1)
gene03 <- sgene("Diameter",     1       ,status = T,start=-40,stop=40,dom=1)
group1 <- rbind(gene01,gene02,gene03)

#positive branch
gene04 <- sgene("X_1X",          1,status = T,start=0,stop=1,dom=50)
gene05 <- sgene("Y_1Y",          1,status = T,start=0,stop=1,dom=50)
gene06 <- sgene("Z_1Z",          1,status = T,start=0,stop=1,dom=50)

#bridge in x
gene07 <- sgene("X_1X",          0,status = T,start=5,stop=9,dom=49)
gene08 <- sgene("Y_1Y",          0,status = T,start=0,stop=1,dom=49)
gene09 <- sgene("Z_1Z",          -1,status = T,start=0,stop=1,dom=49)

#bridge in x
gene10 <- sgene("X_1X",          -1,status = T,start=5,stop=9,dom=48)
gene11 <- sgene("Y_1Y",          -1,status = T,start=0,stop=1,dom=48)
gene12 <- sgene("Z_1Z",          -1,status = T,start=-5,stop=-1,dom=48)

group2 <- rbind(gene04,gene05,gene06,gene07,gene08,gene09,gene10,gene11,gene12)
#negative branch

geneset <- rbind(group1,group2)

pointset <- genetostlfile4   ("~/Desktop/stl/testn.stl",geneset,pos=spos(0, 0,0),runlim=40,verbose = T)
```

If this shape loads correctly in the render engine of choice (blender,cura,slic3r,cura-lulzbot), then we can be satisfied that all other shapes in shapevol will also render (because we are checking all six possible directions)





## plan for 'i' primitive

The 'i' primitive is an idea I had whereby each new edge has  a vertex at the end of it, and both of these have their own  associated stl primitives. In this way, interctions of stl shapes can be avoided because the vertex is filled only once and the edges do not enter the vertex, creating a void in some slicer engines. 

Steps in implementing this will be: 

- implement the vertex primitive only
- implement the edge primitive only
- combine the two
- Allow variability in diameter/volume

## tetrahedra

Need also to think about how to have non-orthoganal shapes - probably read around how this has been attempted before, as part of the literature review. 









